Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-03-16 Codes/PrepareResults.rmd

@@

# Figure (RMSE of EONR vs RMSE of yield for RF abd BRF)

@@

@@

opcp <- tree$cptable[, 1][which.min(tree$cptable[, 4])]
opfit <- prune(tree, opcp)
# fancyRpartPlot(opfit, sub="")
```

>>>>

# Illustrative yield response curve

+ show to *which degree N affects yield* (yield response to N) 
+ show how the relation between yield estimation error and the effect of N on yield increase is ()

## MB yield response to N

```{r}
#/*--------------------------------*/
#' ## Preparation
#/*--------------------------------*/
# --- MB function --- #
gen_yield_MB <- function(ymax, alpha, beta, N) {
  yield <- ymax * (1 - exp(alpha + beta * N))
  return(yield)
}

# --- extract sites that have wide range of true optimal N --- #
get_rows_optN <- function(data, value, range) {
  res <- data %>%
    .[opt_N %between% c(quantile(opt_N, prob = value - range), quantile(opt_N, prob = value + range))
    ] %>%
    .[,group := value]
}

prep_coef_dt <- 
  field_cell_sf %>%
  data.table() %>%
  .[sim==1,.(unique_cell_id, opt_N, alpha, beta, ymax, m_error)]

set.seed(3957)

coef_dt <- 
  lapply(
    c(0.1, 0.5, 0.9), 
    function(x) get_rows_optN(data=prep_coef_dt, value = x, range = .0005)) %>%
  bind_rows() %>%
  # --- select one site from each group --- #
  .[,.SD[sample(.N, size = 1)], by = group]

seq_N <- 
  seq(min(coef_dt[,opt_N]) -50, max(coef_dt[,opt_N])+100, by=1) %>%
  round()

res_y_dt <-  
  coef_dt %>%
  .[rep(1:nrow(.), each = length(seq_N)), ] %>%
  .[, N := seq_N, by = group] %>%
  .[, det_yield := gen_yield_MB(ymax, alpha, beta, N)] %>%
  .[, yield := det_yield*(1 + m_error)] %>%
  .[, opt_N := round(opt_N)] %>%
  melt(id.vars = c("group", "N", "opt_N"), measure.vars = c("det_yield", "yield")) %>%
  .[, variable := case_when(
    variable == "det_yield" ~ "Yield",
    variable == "yield" ~ "Yield including errors"
  )]

#/*--------------------------------*/
#' ## Visualization
#/*--------------------------------*/
sub_dt <- res_y_dt[variable=="Yield" & N == opt_N]

# --- MB function to N --- #
vis_res_y <- ggplot(res_y_dt[variable=="Yield"]) +
  geom_line(aes(x=N, y=value, color=factor(group))) +
  geom_point(data= sub_dt, aes(x=opt_N, y=value)) +
  geom_segment(data = sub_dt, aes(x = N, xend=N, y = value, yend=-Inf), linetype = "dashed") +
  labs(y = "Yield (kg/ha)") +
  labs(x = "N (kg/ha)") +
  theme_dist +
  theme(legend.position = "none")

# ggplot(res_y_dt) +
#   geom_smooth(aes(x=N, y=value, color=factor(group))) +
#   facet_wrap(~variable, ncol = 2) +
#   labs(y = "Yield (kg/ha)") +
#   labs(x = "N (kg/ha)") +
#   theme_dist
```

## MB yield response and estimated yield response by BRF
+ true yield response curves for two sites within a field
+ points true yield evaluated at the N rate that was actually applied at the site
+ RF-estimated yield at all the potential experimental rates including the actual rate that was applied

```{r}
# /*===========================================*/
#'=  Prepare illustrative yield response curve estimated by RF =
# /*===========================================*/

#/*--------------------------------*/
#' ## Preparation
#/*--------------------------------*/
# --- ML functions --- #
library(grf)
source(here("GitControlled/Codes/0_2_functions_main_sim.R"))

# === Data === #
field_subplot_test_dt <- readRDS(here("Shared/Results/for_writing/field_subplot_test_dt.rds"))

train_dt <- 
  field_subplot_sf %>%
  data.table()

test_dt <- 
  field_subplot_test_dt %>%
  data.table()

# /*===== Train ML =====*/
temp_var_ls <- c("alpha", "beta", "ymax")

set.seed(1356)

# RF <- 
#   RF_run(
#     reg_data = train_dt,
#     var_ls = temp_var_ls
#   )

BRF <- 
  BRF_run(
    reg_data = train_dt,
    var_ls = temp_var_ls
  )

# === Yield prediction === #
# --- All N experimental rates --- #
N_rate <- test_dt[,aa_n] %>% unique() %>% sort()
# For illustration purpose, we use site where either 2nd and 4th N rates are actually application
tg_N_rate <- N_rate[c(2,4)]

# --- for true yield response function --- #
seq_N <- 
  seq(min(N_rate) -30, max(N_rate)+30, by=1)
# seq_N <- 
#   seq(min(N_rate), max(N_rate), by=1)

res_true <- 
  test_dt %>%
  .[aa_n %in% tg_N_rate] %>%
  .[rep(1:nrow(.), each = length(seq_N)), ] %>%
  .[, N := seq_N, by = unique_subplot_id] %>%
  .[,det_yield := gen_yield_MB(ymax, alpha, beta, N)] %>%
  .[,yield := det_yield*(1 + m_error)]

# --- for predicted yield at five experimental N rates --- #
res_pred <- 
  test_dt %>%
  .[aa_n %in% tg_N_rate] %>%
  .[rep(1:nrow(.), each = length(N_rate)),] %>%
  .[, N := N_rate, by = unique_subplot_id] %>%
  # .[,pred_y_RF := predict(RF, newdata = .[, c("N", temp_var_ls), with = FALSE])] %>%
  .[,pred_y_BRF := predict(BRF, newdata = .[, c("N", temp_var_ls), with = FALSE])]

# === For visualization === #
set.seed(2032)
low <- 
  res_true[aa_n == tg_N_rate[1]] %>%
  get_rows_optN(data=., value = 0.2, range = .005) %>%
  .[unique_subplot_id %in% sample(unique(unique_subplot_id), size = 1), unique_subplot_id] %>%
  unique()

high <- 
  res_true[aa_n == tg_N_rate[2]] %>%
  get_rows_optN(data=., value = 0.8, range = .005) %>%
  .[unique_subplot_id %in% sample(unique(unique_subplot_id), size = 1), unique_subplot_id] %>%
  unique()

vis_res_true <- res_true[unique_subplot_id %in% c(low, high)]
vis_res_pred <- res_pred[unique_subplot_id %in% c(low, high)]

vis_pred_y_res <- 
  ggplot() +
    # --- true yield response curve --- #
    geom_line(
      data=vis_res_true, aes(x=N, y=yield, group=factor(unique_subplot_id))
    ) +
    # --- Predicted yield points --- #
    geom_point(
      data=vis_res_pred, aes(x=N, y=pred_y_BRF, group=factor(unique_subplot_id)), color="red"
    ) +
    geom_point(data = vis_res_true[aa_n==N], aes(x=aa_n, y= yield), color = 'blue') +
    geom_segment(data = vis_res_true[aa_n==N], aes(x = aa_n, xend=aa_n, y = yield, yend=-Inf), linetype = "dashed") +
    labs(y = "Yield (kg/ha)") +
    labs(x = "N (kg/ha)") +
    theme_dist +
    theme(legend.position = "none")
```

@@

## Figure (RMSE of EONR vs RMSE of yield for RF abd BRF)

@@

## True Yields vs Predicted Yields by models

```{r}
sample_simRes_test_y <- readRDS(here("Shared/Results/for_writing/sample_simRes_test_y.rds"))

rmse <- sample_simRes_test_y %>%
  .[, .(
    rmse_y_det = rmse_general(pred_yield, det_yield),
    rmse_y = rmse_general(pred_yield, yield)
    ), by= .(Method, Model)]

# === Function for RMSE Calculation === #
rmse_general <-function(preds,actual){ 
  sqrt(mean((actual - preds)^2)) %>%
  round(.,1) %>%
  format(, nsmall = 1)

}

rmse <- sample_simRes_test_y %>%
  .[, .(
    rmse_y_det = paste0("RMSE = " ,rmse_general(pred_yield, det_yield)),
    rmse_y = paste0("RMSE = ", rmse_general(pred_yield, yield))
    ), by= .(Method, Model)]

vis_y_pred <- ggplot(sample_simRes_test_y) +
  geom_point(aes(x=det_yield, y=pred_yield), size = 0.5) +
  geom_abline(aes(intercept = 0, slope = 1), color = "red", show.legend = TRUE) +
  geom_text(data=rmse, aes(x=-Inf,y=+Inf,label=rmse_y_det),
              hjust = -0.1,  vjust = 3, size=3.5, family = "Times New Roman") +
  facet_grid(Model ~ Method) +
  guides(
    fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype = guide_legend(keywidth = 3, keyheight = 1),
    colour = guide_legend(keywidth = 3, keyheight = 1)
  ) +
  labs(y = "Predicted Yield (kg/ha)") +
  labs(x = "True Yield without error (kg/ha)") +
  theme_dist
```

@@

opcp <- tree$cptable[, 1][which.min(tree$cptable[, 4])]
opfit <- prune(tree, opcp)
# fancyRpartPlot(opfit, sub="", shadow.col = "")
```

### 2022-03-16 Writing/manuscript_cea.rmd

The first term on the right-hand side of Eq. \@ref(eq:eqn3) is the deterministic yield component. The variable $N$ represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn3), $\alpha$, $\beta$, and $ymax$ are spatially varying parameters that govern how yield level is determined in conjunction with the applied N rate. They reflect local cell-level production conditions, such as soil characteristics and topography. For example, $\alpha$, which represents what the yield would be when no nitrogen is applied in conjunction with $ymax$, may reflect soil organic matter and residual nitrogen from previous year's production. $ymax$, which determines the maximum yield attainable, may include other micro nutrients like phosphorus and soil depth. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. The error term of the yield $\varepsilon$ represents the composite of all the unobserved factors that affect crop yield beyond $\alpha$, $\beta$, $ymax$ and $N$. Each cell $(i,j)$'s "true" optimal N application rate, denoted $N^{opt}_{i,j}$, was calculated by solving a profit maximization problem by cell assuming that the prices of corn and N fertilizer were $\$0.138/kg$ and $\$1.323/kg$, respectively. These prices are used consistently for our simualtions. See Appendix C.1 for further details on how the paremeters were generated and sample maps of the parameters and true economically optimal N rates.

@@

The accuracy of EONR estimation was judged based on RMSE ($kg/ha$) of the EONR estimation against the true EONR at the subplot level. Using the EONR estimates, we also calculated profit-deficits ($\$/ha$) (i.e., $\hat{\pi}_{def}$) relative to the true maximum profit at the subplot level, defined as profit under the true yield response functions evaluated at $N^{opt}_{i}$. For RF, BRF, and CNN, the accuracy of yield prediction was judged based on the RMSE ($kg/ha$) of the yield prediction against the true yields at the subplot level.

>>>>

The first term on the right-hand side of Eq. \@ref(eq:eqn3) is the deterministic yield component. The variable $N$ represents the nitrogen fertilizer application rate (kg/ha) which the trial designs varied by plot. Also in Eq. \@ref(eq:eqn3), $\alpha$, $\beta$, and $ymax$ are spatially varying parameters that govern how yield level is determined in conjunction with the applied N rate. They reflect local cell-level production conditions, such as soil characteristics and topography. For example, $\alpha$, which represents what the yield would be when no nitrogen is applied in conjunction with $ymax$, may reflect soil organic matter and residual nitrogen from previous year's production. $ymax$, which determines the maximum yield attainable, may include other micro nutrients like phosphorus and soil depth. $\beta$ can be interpreted as reflecting how efficiently the crop uses N to create grain mass, and might depend on soil properties that affect the availability of applied N fertilizer to the crop, say due to N immobilization, leaching, denitrification, and volatilization [@johnson2005nitrogen; @alva2006nitrogen]. The error term of the yield $\varepsilon$ represents the composite of all the unobserved factors that affect crop yield beyond $\alpha$, $\beta$, $ymax$ and $N$. Each cell $(i,j)$'s "true" optimal N application rate, denoted $N^{opt}_{i,j}$, was calculated by solving a profit maximization problem by cell assuming that the prices of corn and N fertilizer were $\$0.138/kg$ and $\$1.323/kg$, respectively. These prices are used consistently for our simulations. See Appendix C.1 for further details on how the paremeters were generated and sample maps of the parameters and true economically optimal N rates.

@@

The accuracy of EONR estimation was judged based on RMSE ($kg/ha$) of the EONR estimation against the true EONR at the subplot level:

\begin{equation}
\mbox{RMSE of EONR estimation} = \sqrt{\frac{1}{1536}\sum_{i=1}^{1536}(N^{opt}_{i} - \hat{N}^{opt}_{i})^2}. (\#eq:eqn7)
\end{equation}

Using the EONR estimates, we also calculated profit-deficits ($\$/ha$) (i.e., $\hat{\pi}_{def}$) relative to the true maximum profit at the subplot level, defined as profit under the true yield response functions evaluated at $N^{opt}_{i}$. For RF, BRF, and CNN, the accuracy of yield prediction was judged based on the RMSE ($kg/ha$) of the yield prediction against the true yields at the subplot level:

\begin{equation}
\mbox{RMSE of yield prediction} = \sqrt{\frac{1}{1536}\sum_{i=1}^{1536}(y_{i} - \hat{y}_{i})^2}. (\#eq:eqn8)
\end{equation}
